---
title: "Local Least Squares Imputation"
output:
  pdf_document: default
  word_document: default
date: "`r Sys.Date()`"
---
#Local Least Squares Imputation Method

The core tidyverse includes the packages that are likely to use in everyday 
data analysis.
```{r}
install.packages("tidyverse")
```
To retrive the packages stored in the directory execute library function
```{r}
library("tidyverse")
```

To visualize dataset and getting a brief look into missing data pattern 'visdat' 
package should be installed. Further pacakages are retrived using library
```{r}
install.packages("visdat")
library("visdat")
```

To read excel files "readxl" package is installed as by default this is missing
in tidyverse packages
```{r}
install.packages("readxl")
library("readxl")
```

Now, dataset should be imported to analysis the missing pattern and perform 
imputation. Before importing the dataset set the working directory with respect
to dataset location using session tile. In the below code col_names is set
false to avoid the first row as column name.
```{r}
inc_data <- read_excel("Ionosphere_NE_10.xlsx", col_names  = FALSE)
```

Change of column names.
```{r}
inc_data <- setNames(data.frame(inc_data), c('col1','col2','col3','col4',
                                             'col5','col6','col7','col8','col9','col10','col11','col12','col13','col14','col15',
                                             'col16','col17','col18','col19','col20','col21','col22','col23','col24','col25','col26',
                                             'col27','col28','col29','col30','col31','col32','col33','col34'))
```

Display colnames
```{r}
colnames(inc_data)
```


#Analyzing dataset using below commands:

Overview of the dataset.
```{r}
summary(inc_data)
sum(is.na(inc_data)) #extractsno of missing values.
```

structure of the dataset
```{r}
str(inc_data)
```

Boxplot for individual columns 
```{r}
boxplot(inc_data)
```

Visualisation of missing data using 'visdat'
```{r}
install.packages("visdat")
library("visdat")
vis_miss(inc_data, sort_miss = TRUE) # TRUE is set descending order/high percent missing first
vis_dat(inc_data)
```

pcaMethods package installation
```{r}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("pcaMethods")
```

Extracting library functions in pcaMethods package
```{r}
library(pcaMethods)
```

Using llsImpute function for imputation.
In this function parameters are k, correlation, and allVariables
k is selected based on no: of columns, missing pattern, no: of missing values.
correlation is selected as per the reference paper.
allVariables is 'TRUE' for complete variables
```{r}
impu_data <- llsImpute(inc_data, k=5, correlation = "pearson", allVariables = TRUE)
```

The estimated values gets stored in completeObs variable this can be extracted
using below code chunck
```{r}
est_data <- completeObs(impu_data)
est_data
```

Formation of data frame for the given array of values
```{r}
est_data <- data.frame(est_data)
```

Checking 'NA' values in estimated dataset
```{r}
sum(is.na(est_data))
```
Viz for estimated data
```{r}
vis_miss(est_data)
```

Importing original dataset to find NRMS
```{r}
ori_data <- read_excel("Ionosphere.xlsx", col_names  = FALSE)
```

Change of column name
```{r}
ori_data <- setNames(data.frame(ori_data), c('col1','col2','col3','col4',
                                             'col5','col6','col7','col8','col9','col10','col11','col12','col13','col14','col15',
                                             'col16','col17','col18','col19','col20','col21','col22','col23','col24','col25','col26',
                                             'col27','col28','col29','col30','col31','col32','col33','col34'))
```


To check datatype and change if required. 
```{r}
typeof(ori_data)
ori_data <- data.matrix(ori_data, rownames.force = NA)
```

Now, using frobenius method to find the normalized value to calculate NRMS using 'norm' function
compute norm of a matrix 
NRMS = || Xest - Xori||/||Xori||
finding numerator || Xest - Xori|| using norm function, type frobenius form
```{r}
diff_data <- (est_data - ori_data)
typeof(diff_data)
```
Convert list to double for normalization 
```{r}
diff_data <- data.matrix(diff_data, rownames.force = NA)
typeof(diff_data)
```

Top 6 rows for diff_data
```{r}
head(diff_data)
```

Numerator normalization using 'norm()' for NRMS
```{r}
num <- norm(diff_data, type = c("F"))
num
```
Denominator normalization using 'norm()' for NRMS
```{r}
den <- norm(ori_data, type = c("F"))
den
```
Calculating NRMS
```{r}
NRMS <- num/den
NRMS
```